A GPU Implementation of Capon for an Active Sonar System

### 2012-05-03 ECUA2012/gpu_capon_ecua2012.tex
la2pdf commit - 2012:05:03	15:00:49	(Thu 3.May)

Consider an $M$ element phased array. Let the sample recorded by the $m$'th channel at time instant $n$ be represented as $x_m[n]$, and the accompanying weights and delays be  $w_m[n]$ and $\Delta_m$. The beamformer output $z[n]$ is then given as
\begin{align}
z[n] = \w\H[n]\X[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n-\Delta_0]\\x_1[n-\Delta_1]\\\vdots\\x_{M-1}[n-\Delta_{M-1}]}.
\end{align}
The \gls{MVDR} beamformer \cite{Capon1969} finds the set of weights that minimizes the output power, $\min E\{|z[n]|^2\}$, while ensuring unity gain in the look direction. Assuming predelayed data, the solution to this optimization problem is given as
\begin{gather}
\vec w[n] = \frac{\eRi[n]\1}{\1\T\eRi[n]\1} = \frac{\vec\beta[n]}{\1\T\vec\beta[n]},
\end{gather}
where $\1$ is a row vector of ones, and $\eR$ is an estimate of the spatial covariance matrix. To ensure robustness in an active sonar setting, we perform subarray averaging in order to decorrelate noise from the signal, and some degree of temporal averaging to retain \gls{DAS}-like speckle statistics. The $(i,j)$'th element of the sample covariance matrix, $\hat r_{ij}$, is then given as
\begin{gather}
r_{ij}[n] = \frac{1}{(M-L+1)(2N_{\text{avg}}+1)}\sumb{k=0}{M-L} \sumb{n'=n-N_{\text{avg}}}{n+N_{\text{avg}}} x_{i+k}[n']\,x_{j+k}[n'],
\end{gather}
where $r_{ij}$ is the $(i,j)$'th element of the covariance matrix, $L$ is the subarray size, and $2N_{\text{avg}}+1$ is the number of temporal samples to perform averaging over. Finally, to make sure that $\eR$ is well conditioned, we add a fraction $d$ of the total power of $\eR$ to its diagonal\cite{Synnevag2007},
\begin{align}
\eR = \eR + \I \cdot d \cdot \tr\{\eR[n]\}.
\end{align}

@@

>>>>

Consider an $M$ element phased array. Let the sample recorded by the $m$'th channel at time instant $n$ be represented as $x_m[n]$, and the accompanying weights and delays be  $w_m[n]$ and $\Delta_m[n]$. The beamformer output $z[n]$ is then given as
\begin{align}
z[n] = \w\H[n]\X[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n-\Delta_0]\\x_1[n-\Delta_1]\\\vdots\\x_{M-1}[n-\Delta_{M-1}]}.
\end{align}
The \gls{MVDR} beamformer \cite{Capon1969} finds the set of weights that minimizes the output power, $\min E\{|z[n]|^2\}$, while ensuring unity gain in the look direction. The solution to this optimization problem is given as
\begin{gather}
\vec w[n] = \frac{\Ri[n]\1}{\1\T\Ri[n]\1},
\end{gather}
where $\1$ is a row vector of ones, and $\R$ is the spatial covariance matrix. 
\begin{gather}
R[n] = \frac{1}{2N_\text{avg}+1} \sumb{n'=n-N_\text{avg}}{n+N_\text{avg}} \x[n']\x\H['n]
\end{gather}

To ensure robustness in an active sonar setting, we perform subarray averaging in order to decorrelate noise from the signal, and some degree of temporal averaging to retain \gls{DAS}-like speckle statistics. The $(i,j)$'th element of the sample covariance matrix, $\hat r_{ij}$, is then given as
\begin{gather}
r_{ij}[n] = \frac{1}{(M-L+1)(2N_{\text{avg}}+1)}\sumb{k=0}{M-L} \sumb{n'=n-N_{\text{avg}}}{n+N_{\text{avg}}} x_{i+k}[n']\,x_{j+k}[n'],
\end{gather}
where $r_{ij}$ is the $(i,j)$'th element of the covariance matrix, $L$ is the subarray size, and $2N_{\text{avg}}+1$ is the number of temporal samples to perform averaging over. Finally, to make sure that $\eR$ is well conditioned, we add a fraction $d$ of the total power of $\eR$ to its diagonal\cite{Synnevag2007},
\begin{align}
\eR = \eR + \I \cdot d \cdot \tr\{\eR[n]\}.
\end{align}

@@

\begin{gather}
\vec w[n] = \frac{\Ri[n]\1}{\1\T\Ri[n]\1} = \frac{\vec\beta[n]}{\1\T\vec\beta[n]},
\end{gather}
